---
title: "scenario_2_r_markdown"
author: "Grace E. Seo"
date: "2/1/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Scenario 2

In Scenario 1, you performed some analysis and created a scatter plot of Pokemon bioinformatics class using student metadata.

To provide the result to your supervisor, you need to create a PDF output using R markdown.


Before you begin, let's learn how R markdown works briefly.

If you want to know more about R markdown, here is a great resource:
https://bookdown.org/yihui/rmarkdown-cookbook/


In R markdown, you can create a Bookmark/table of contents by using #.

For example:

# Header 1
## Header 2
### Header 3




### Bullet points

If you want to include bullet points, use "- " in front of your sentence. 

- Like this! Make sure you put a space before the dash!

- If you want to leave some blank lines between sentences, include a white space between two sentences by pressing "Enter".



### Include a timestamp on this document

To know when this PDF file was generated, include a timestamp.

date: "Last compiled on `r format(Sys.time(), '%d %B, %Y, at %H:%M:%S')`"



Now, let's include all the codes from scenario 1 as well include outputs directly in this PDF file.

\newpage


## 1. Load Pokemon class metadata

GitHub page: https://github.com/MMID-coding-workshop/2022-02-02-Introduction-to-R

Raw Pokemon class metadata: https://github.com/MMID-coding-workshop/2022-02-02-Introduction-to-R/blob/main/class_metadata.csv

```{r }
## Raw csv file: https://github.com/MMID-coding-workshop/2022-02-02-Introduction-to-R/blob/main/class_metadata.csv
## Two methods to load dataset: using a URL or directly from your folder.

# CODE: OBJECTNAME <- read.csv("FOLDERPATH/FILE.csv", header = TRUE)
# CODE: OBJECTNAME <- read.csv("URL/FILE.csv", header = TRUE)

#metadata <- read.csv("C:/Users/USERNAME/Desktop/intro_to_r/class_metadata.csv", header = TRUE)

metadata <- read.csv("https://raw.githubusercontent.com/MMID-coding-workshop/2022-02-02-Introduction-to-R/main/class_metadata.csv", header = TRUE)

## NOTE: now you should see that "metadata" object is available in your "Environment" pane

```



### View dataset

```{r }

# CODE: View("OBJECTNAME") 

View(metadata)

## NOTE: new window will pop-up beside this script. You can scroll through the content.

metadata # this will print data into console.

```

\newpage

## 2. Install and load necessary package for your analysis if not already installed

```{r }
## For simple analyses, we will install tidyverse package available in CRAN
## NOTE: tidyverse package includes tidyr, dplyr, ggplot2 and many more!
# CODE: install.packages("PACKAGENAME")

#install.packages("tidyverse") 

## NOTE: if you already have tidyverse installed, comment out the above code so you don't repeat installation.


## Load installed package == library
# CODE: library(PACKAGENAME)

library(tidyverse)

```


\newpage

## 3. Determine how many students are in your class
HINT: Row number matches student number

```{r }
#CODE: nrow(OBJECTNAME) # counts how many rows are present in a dataframe

nrow(metadata)

```

ANSWER: 20 students


## 4. Determine how many columns are present in your metadata

```{r }
# CODE: ncol(OBJECTNAME) # counts how many columns are present in a dataframe

ncol(metadata)

```

ANSWER: 9



## 5. Show the list of column headers in the metadata

```{r }
# CODE: colnames(OBJECTNAME) # prints out list of column headers

colnames(metadata)

```

ANSWER:
[1] "Number"         "Student_name"   "Student_number" "Age"            "Year_of_study"  "Phone"         
[7] "Address"        "Address_city"   "Major"   


\newpage

## 6. Print out row contents of only column "Student_name"

```{r }
# CODE: OBJECTNAME$COLNAME

metadata$Student_name

```

ANSWER:
[1] "Mew"            "Garchomp"       "Kanto Starters" "Metagross"      "Zoroark"        "Tyranitar"     
[7] "Eevee"          "Snorlax"        "Volcarona"      "Mewtwo"         "Charizard"      "Milotic"       
[13] "Kyurem"         "Haxorus"        "Arceus"         "Ninetales"      "Espeon"         "Golurk"        
[19] "Lapras"         "Pikachu" 



## 7. Determine average age of all students in this class

```{r }
# CODE: mean(OBJECTNAME$COLNAME)

mean(metadata$Age)

```

ANSWER: 24



## 8. Determine the Pikachu's student number

```{r }
# CODE: OBJECTNAME$COLNAME_CONTAIN_VALUE_YOU_WANT[OBJECTNAME$COLNAME_YOUR_SEARCH == "NAME"]

metadata$Student_number[metadata$Student_name=="Pikachu"]

```

ANSWER: 7094583


\newpage

## 9. Scatter plot by student's Year_of_study vs. Age and create a figure legend by student's study Major

```{r, fig.cap = "Student's Year of Study by Age and Major represented as simple scatter plot."}
# CODE: NEWOBJECT <- ggplot(data = OBJECTNAME, aes(x = X_VALUE, y = Y_VALUE, colour = GROUP)) +
#                    geom_point()

# CODE: NEWOBJECT # to view the content


plot <- ggplot(data = metadata, aes(x = Year_of_study, y = Age, colour = Major)) +
        geom_point()

plot # view the plot

```


\newpage

## 10. Scatter plot is too small to see. 
Increase the size of each scatter point, remove grey background and rename the x-axis to "Year of Study". 


```{r, fig.cap = "Student's Year of Study by Age and Major represented as scatter plot."}
plot_2 <- ggplot(data = metadata, aes(x = Year_of_study, y = Age, colour = Major)) +
    geom_point(size = 5) +
    xlab("Year of Study") +
    theme_bw()

plot_2 # view the plot

```



This is it! Now click "Knit" at top on this R markdown --> "Knit to PDF" or "Knit to HTML" and you'll get a file output.


The end of scenario 2


\newpage


### Upcoming workshops

Now try this with your own code and plots.
In upcoming R workshops, you'll be able to learn in-depth on how to do:

- more sophisticated data transformation
- include more options and make fancier graphs in ggplot, ggtree, etc.
- perform RNA-seq data analyses
- and build machine learning models!



## References

If you want to practice R markdown in detail, here are helpful resources:

1. Yihui Xie, Christophe Dervieux, Emily Riederer. 2022. R Markdown Cookbook. Available at https://bookdown.org/yihui/rmarkdown-cookbook/.

2. Chester Ismay and Patrick C. Kennedy. 2021. Getting Used to R, RStudio, and R Markdown. https://ismayc.github.io/rbasics-book/index.html.